Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-05-03 paper/paper.tex
interm2

\subsubsection{Optimal control via stochastic gradient descent in the DMN}
Efficient learning of the entire set of model parameters can effectively be achieved
via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
local minima of differentiable functions. Stochastic here that the true gradient is estimated from batches of training examples samples from the replay memory, the HC

@@

% \paragraph{Parameter update.}
\begin{equation}
  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}].
  \label{eq:oracle}
\end{equation}
Thus next action is taken to drive reward prediction errors \textemdash also known as \textit{regret}\textemdash percolate
from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
the thus generated approximating sequence of
$Q$-value functions provably
% \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.

One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
the above framework using a linear value approximator with feature encoding
$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
$\mathcal S \times \mathcal A$, and so
$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
Whence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
$$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
the well-known TD update rule \citep{sutton1998reinforcement}.

\subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
In RL, Monte-Carlo simulation can be used to update the agent's belief state
\citep{silver2010monte}.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.
Monte-Carlo simulation provides a simple method for evaluating the value of a state.
They provide an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.

@@

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
  activity when humans engage in cognitive processes that are detached from
  current sensory environment
  \citep{kenet2003spontaneously, fiser2004small, bzdok2013segregation}
  and it was proposed to be situated at the top of the brain network hierarchy
  \citep{carhart2010default, margulies2016situating}.
  Its putative involvement in thinking about the past,
  the future, and hypothetical experiences
  appears to tie in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy encapsulates the repertoire of possible actions
    on the world given a current state.
    The policy matrix encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
% information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  neurobiological implementation of reinforcement learning.}
  Overview of how the constituent nodes of the DMN
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

The HC may provide perturbed action-transition-state-reward samples
  as batches of imagined, hypothesized, and recalled experience.
  The small variations in these experience samplings allow searching
  a larger space of model parameters and possible experiences.
  Taken to its extreme, stochastic recombination of experience
  building blocks can further optimize the behavior of the RL agent
  by model learning from scenarios in the environment that the agent might
  only very rarely or never encounter.
  An explanation is thus offered for experiencing seemingly familiar situations that
  a human has however never actually encountered (i.e., d\'{e}j\`{a} vu effect).
  While such a situation may not have been experienced in the physical world,
  the DMN may have previously stochastically generated, evaluated, and adapted to
  such a randomly synthesized situation.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep),
  mental scene construction allows for pseudo-experiencing possible
  future scenarios and action outcomes.
  Our formal account of DMN function
  thus acknowledges the unavoidable stochasticity of
  computation in neural systems \citep{faisal2008noise}.

>>>>

\subsubsection{Optimal control via stochastic gradient descent in the DMN}
Efficient learning of the entire set of model parameters can effectively be achieved
via stochastic \textit{gradient descent}, a first-order\footnote{ i.e using only differential information up to order 1.} algorithm for finding
local minima of differentiable functions.
Stochastic here means that the true gradient is estimated from batches of training examples samples from the replay memory, the HC

@@

% \paragraph{Parameter update.}
\begin{equation}
  \delta \theta_{k} \propto -\nabla_{\theta_{k}}\mathcal L(\theta_{k})
  = -\mathbb E_{(s, a, r, s') \sim \mathcal D}[\underbrace{(\tilde{Q}(s, a|\theta_{k}) - y_k)}_{\text{prediction error}}
    \underbrace{\nabla_{\theta_{k}}\tilde{Q}(s, a|\theta_{k})}_{\text{aversion}}].
  \label{eq:oracle}
\end{equation}
Thus, the next action is taken to drive reward prediction errors
(or \textit{regret}) to
percolate
from lower to higher processing layers to modulate the choice of future actions .Over time (i.e., in the limit $k \rightarrow \infty$),
the thus generated approximating sequence of
$Q$-value functions provably
% \footnote{The proof is standard and relies on the Banach Fixed-Point Theorem applied to the Bellman transform.}
$\tilde{Q}(.,.|\theta_0) \rightarrow \tilde{Q}(.,.|\theta_1) \rightarrow \tilde{Q}(.,.|\theta_2) \rightarrow \ldots$
are attracted and absorbed by the optimal $Q$-value function $Q^*$ defined implicitly by the Bellman equation \eqref{eq:bellman}.

One should note that most classical RL algorithms, including \textit{Temporal Difference (TD)}
learning \citep{sutton1998reinforcement}, REINFORCE \citep{williams1992}, and SARSA can be
cast in the general variational framework \eqref{eq:oracle}. For instance, TD corresponds to
the above framework using a linear value approximator with feature encoding
$\phi(s,a) = \delta_{(s,a)} =  \text{ point mass at }(s,a)$ on the grid
$\mathcal S \times \mathcal A$, and so
$$\nabla_{\theta}\tilde{Q}(s, a|\theta) = \phi(s, a) = \delta_{(s,a)}.$$
Hence, the gradient update due to the sample $(s,a,r,s') \in \mathcal D$ is
$$\tilde{Q}(s, a) \leftarrow \tilde{Q}(s, a) + \alpha(y-\tilde{Q}(s, a)),$$
the well-known TD update rule \citep{sutton1998reinforcement}.

\subsubsection{Does the hippocampus subserve Monte-Carlo sampling?}
In RL, Monte-Carlo simulation is a common means to update the agent's belief state
\citep{silver2010monte}.
Monte-Carlo simulation provides a simple method for evaluating the value of a state.
They provide an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

@@

\subsection{Putting everything together}
The DMN is today known to consistently increase in neural
  activity when humans engage in cognitive processes that are detached from
  the current sensory environment
  \citep{kenet2003spontaneously, fiser2004small, bzdok2013segregation}
  and this network was proposed to be situated
  at the top of the brain network hierarchy
  \citep{carhart2010default, margulies2016situating}.
  Its putative involvement in thinking about the past,
  the future, and hypothetical experiences
  appears to tie in with the implicit computation of
  action and state cascades as a function of what happened in the past.
  A policy encapsulates the repertoire of possible actions
    on the world given a current state.
    The policy matrix encodes the probabilites of
    choosing actions to be executed in a certain situation.
% Interactions between external and internal sources of
% information are probably commonplace during perception \citep{gershman2010learning}
The DMN may subserve
  constant exploration of possible future actions and estimaton of their
  cumulative reward outcomes. Implicit computation of future choices
  provides an explanation for the
  evolutionary emergence and practical usefulness of
  mind-wandering at day-time and dreams during sleep
  in humans.
\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  neurobiological implementation of reinforcement learning.}
  Overview of how the constituent nodes of the DMN
  map onto computational components necessary for an RL agent.}
  \label{fig:rl_process_chart}
\end{figure}

The HC may provide perturbed action-transition-state-reward samples
  as batches of pseudo-experience
  (i.e., imagined, hypothesized, and recalled mental scenarios).
  The small variations in these experience samplings allow searching
  a larger space of model parameters and possible experiences.
  Taken to its extreme, stochastic recombination of experience
  building blocks can further optimize the behavior of the RL agent
  by model learning from scenarios in the environment that the agent might
  only very rarely or never encounter.
  An explanation is thus offered for experiencing seemingly familiar situations that
  a human has however never actually encountered (i.e., d\'{e}j\`{a} vu effect).
  While such a situation may not have been experienced in the physical world,
  the DMN may have previously stochastically generated, evaluated, and adapted to
  such a randomly synthesized situation.
  In the absence of environmental input and feedback
  (e.g., mind-wandering or sleep),
  mental scene construction allows for pseudo-experiencing possible
  future scenarios and action outcomes.
  Our formal account of DMN function
  thus acknowledges the unavoidable stochasticity of
  computation in neural systems \citep{faisal2008noise}.
